---
title: "Titanic Kaggle Competition"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('ggplot2') # visualization
library('dplyr') # data manipulation
library('mice') # imputation
library('randomForest') # classification algorithm
```

## Exploring Machine Learning with Kaggle's Titanic Competition

Here's my code for Kaggle's introductory machine learning competition. Let's see who is going to live and who is going down with the ship


```{r, echo = FALSE}


train <- read.csv("./data/train.csv", stringsAsFactors = FALSE)
test <- read.csv("./data/test.csv", stringsAsFactors = FALSE)

full <- bind_rows(train, test)
```

## Feature Engineering

Here is the code to grab the title, Mr., Mrs, etc., from the name variable and create a new variable called title. I also am creating a variable called Fsize that is the family size, including the person. I also standardize Mlle, Ms, Mme and "rare titles."

```{r}
library(knitr)
full <- full %>%
  mutate(Title = (gsub('(.*, )|(\\..*)', '', Name)),
         Surname = (gsub('(,)(.)*', '', Name)),
         Fsize = (SibSp + Parch + 1),
         Family = paste(Surname, Fsize, sep = '_')
         )

# Titles with very low cell counts to be combined to "rare" level
rare_title <- c('Dona', 'Lady', 'the Countess','Capt', 'Col', 'Don', 
                'Dr', 'Major', 'Rev', 'Sir', 'Jonkheer')

# Also reassign mlle, ms, and mme accordingly
full$Title[full$Title == 'Mlle']        <- 'Miss' 
full$Title[full$Title == 'Ms']          <- 'Miss'
full$Title[full$Title == 'Mme']         <- 'Mrs' 
full$Title[full$Title %in% rare_title]  <- 'Rare Title'

# Show title counts by sex again
table(full$Sex, full$Title)
```


## Here's a bar chart to show how family size affects survival:

Notice the difference in survival between singletons, small (2-4 members), and larger, >4, family sizes.

```{r}
full %>%
  filter(!is.na(Survived)) %>%
  ggplot(mapping = aes(x = Fsize, fill = factor(Survived))) +
    geom_bar(stat = "count", position = "dodge") +
    scale_x_continuous(breaks = c(1:11)) +
    labs(x = "Family Size")
```

## Combining family sizes

Since being a singleton or a large family size >4 is bad for survival and a small family size is good for surviving, let's create a variable that puts people in singleton, small, or large family categories.

```{r}
# Discretize family size
full$FsizeD[full$Fsize == 1] <- 'singleton'
full$FsizeD[full$Fsize < 5 & full$Fsize > 1] <- 'small'
full$FsizeD[full$Fsize > 4] <- 'large'

# Show family size by survival using a mosaic plot
mosaicplot(table(full$FsizeD, full$Survived), main='Family Size by Survival', shade=TRUE)
```
